--- 
title: "Machine learning orientation"
author: "Uwe Sterr"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
bibliography: [/Users/uwesterr/CloudProjectsUnderWork/ProjectsUnderWork/bibliographyGeneral.bib]
biblio-style: apalike
link-citations: yes
description: "A guide to machine learning"
---


```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE, cache = TRUE)

```

# Introduction

This site gives on orientation to approach machine learning from the question **shall we do ML to coding examples** for ML algorithms. The site is organized in a way that it covers first the basics in the main chapters and goes into detail in the sub chapters.  
The site is a living document, i.e. there are TBDs and TBCs where work is ongoing.

To find help on how to navigate the site please press the "i" icon on top of the page.

```{block2 echo=TRUE, type='rmdtip'}
**Example topics:**
  
- Machine learning: Shall we? in chapter \@ref(whatML)
- ML examples in chapter \@ref(mlAplicationExamples)
- Real world example on signal detection in chapter \@ref(RealWorld) 
- Further examples from Kaggle  in chapter \@ref(KaggleExamples) 
- Explainable machine learing in chapter \@ref(ExplainableMl)
- Overview online courses in chapter \@ref(MlMooc)
- Overview online information resources in chapter \@ref(MlResources)
- Analysis of the 2019 Kaggle member survey in chapter \@ref(KaggleSurvey)
- How to get quickly started with cloud based machine learning platforms in chapter \@ref(CloudBasedMl) 


```

```{r eval=FALSE, include=FALSE}
<embed src="../MachineLearningOrientierung.pdf" width="800" height="875" 
 type="application/pdf">
```






<video width="80%" loop controls>
  <source src="images/VaihingenTubMovieV3WithSteeringAndSalient.mp4" type="video/mp4">
  Your browser does not support the video tag.
</video>





